---
title: "p8105_hw5_ps3070"
output: html_document
---


```{r}
library(tidyverse)
```

## Problem 1
```{r iris data}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))

is.list(iris_with_missing) #confirmed that it is a list
```

There are two cases to address:

For numeric variables, you should fill in missing values with the mean of non-missing values

For character variables, you should fill in missing values with "virginica"

Write a function that takes a vector as an argument; replaces missing values using the rules defined above; and returns the resulting vector. Apply this function to the columns of iris_with_missing using a map statement.


```{r}
replace_iris_missing = function(a) {
  if (is.numeric(a)) {
    b = mean(a, na.rm = TRUE) #creating new variable for means
    a = replace(a, is.na(a), b) #replacing na values in a with means from b
  } else if (is.character(a)) {
    a = replace(a, is.na(a), "virginica") #replacing character na values
  }
}

output = map_df(iris_with_missing, replace_iris_missing) #maps function to each vector 
output
```

## Problem 2

This dataset contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

Start with a dataframe containing all file names; the list.files function will help
Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe
Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary
Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

## Problem 3
